Forecasting Influenza Hospitalizations Using a Bayesian Hierarchical Nonlinear Model with Discrepancy

### 2024-12-10 nonlinear_forecast.tex
Push some changes

\subsection{Influenza-like illness data}
The US Outpatient Influenza-like Illness Surveillance Network (ILINet) collects information on respiratory illness from outpatient visits to health care providers. Over 3,400 outpatient health care providers in all 50 US states, PR, DC, and the US Virgin Islands report each week the total number of outpatient visits along with the number of ILI cases. An ILI case is defined as a ``fever (temperature of $100^{\circ}$F[$37.8^{\circ}$C] or greater) and a cough and/or a sore throat." Prior to the 2021 season, the definition included ``without a known cause other than influenza" \cite[]{cdcfluview2023}. Because other illnesses such as COVID-19, RSV, and the common cold may induce similar respiratory symptoms, ILI may include patients infected with some disease other than influenza. To know whether or not a sick patient is infected with influenza would require a laboratory test.

@@

\begin{figure}
    \centering
    \includegraphics[scale=.5]{crps_by_season.png}
    \caption{Boxplots of continuous ranked probability score (CRPS) for the four ILI models over all weeks and horizons in the simulation study faceted by season and including seasons 2010-2022, excluding 2020}
    \label{fig:crps_by_season}
\end{figure}

@@

Figure \ref{fig:us_lwis} shows model performance by LWIS for each week of the season for all 24 models of US hospitalizations. LWIS scores are grouped by ILI model, hospitalization model distribution, and by the linear or quadratic modeling. Here, a darker value represents a larger LWIS and worse forecast compared to lighter values. The weeks around the holiday week 22 are highlighted by a red band. In most cases, there appears to be a turning point in performance at or near week 22. The models which include discrepancy appear to forecast better around week 22, though they may not outperform the models without discrepancy through the whole season. 
Such is the case for the NORM and LST quadratic models and the linear LNORM models. It's also notable that for the SIR model, including discrepancy appears to make for poorer forecasts after week 22 whereas for ASG, including discrepancy appears to improve forecasts.

@@

    % \centering
\includegraphics[scale = .5]{lwis_us_full_season.png}
\caption{Each plot shows the log weighted interval score (LWIS, lighter is better) for every week of the 2023 flu season with scores for models with (bottom) and without (top) including discrepancy. Facets distinguish hospitalization distribution family and ILI as a linear or quadratic predictor (column facets, quadratic indicated with 2) with asymmetric Gaussian or SIR ILI models (row facets).}
    \label{fig:us_lwis}
\end{figure}

>>>>

\subsection{Influenza-like illness data}
The US Outpatient Influenza-like Illness Surveillance Network (ILINet) collects information on respiratory illness from outpatient visits to health care providers. Over 3,400 outpatient health care providers in all 50 US states, PR, DC, and the US Virgin Islands report each week the total number of outpatient visits along with the number of ILI cases. An ILI case is defined as a ``fever (temperature of $100^{\circ}$F[$37.8^{\circ}$C] or greater) and a cough and/or a sore throat.'' Prior to the 2021 season, the definition included ``without a known cause other than influenza'' \cite[]{cdcfluview2023}. Because other illnesses such as COVID-19, RSV, and the common cold may induce similar respiratory symptoms, ILI may include patients infected with some disease other than influenza. To know whether or not a sick patient is infected with influenza would require a laboratory test.

@@

\begin{figure}
    \centering
    \includegraphics[scale=.5]{crps_by_season.png}
    \caption{Boxplots of continuous ranked probability score (CRPS) for the four ILI models over all weeks and horizons in the simulation study faceted by season and including seasons 2010-2022, excluding 2020.}
    \label{fig:crps_by_season}
\end{figure}

@@

Figure \ref{fig:us_lwis} shows model performance by LWIS for each week of the season for all 24 models of US hospitalizations. LWIS scores are grouped by ILI model, hospitalization model distribution, and by the linear or quadratic modeling. Here, a darker value represents a larger LWIS and worse forecast compared to lighter values.  In most cases, there appears to be a turning point in performance at or near week 22. The models which include discrepancy appear to forecast better around week 22, though they may not outperform the models without discrepancy through the whole season. 
Such is the case for the NORM and LST quadratic models and the linear LNORM models. It's also notable that for the SIR model, including discrepancy appears to make for poorer forecasts after week 22 whereas for ASG, including discrepancy appears to improve forecasts.

@@

    % \centering
\includegraphics[scale = .5]{lwis_us_full_season}
\caption{Each plot shows the log weighted interval score (LWIS, lighter is better) for every week of the 2023 flu season with scores for models with (bottom) and without (top) including discrepancy. Facets distinguish hospitalization distribution family and ILI as a linear or quadratic predictor (column facets, quadratic indicated with 2) with asymmetric Gaussian or SIR ILI models (row facets). The weeks around the holiday week 22 are highlighted by a red band.}
    \label{fig:us_lwis}
\end{figure}
